Prestige as a Driving Force in Cultural Transmission

### 2022-06-06 ms/PrestigeBias.tex
update

\paragraph{Influence bias.}
Today, social media provides an easy way to estimate the social and cultural influence individuals have over others, and therefore may have an effect on decision making. Online social networks such as \textit{Facebook} or \textit{Instagram} are known to affect the influence of individuals \citep{social_influence,influence_analysis,social_media}, and specific marketing practices were invented to capitalize on this effect \citep{facebook_marketing}.
Here, we model such influence as an indirect bias in cultural transmission, in which the choice of a role-model depends on the choices made by other individuals that have already chosen a role-model. 
Other than its relative ease of evaluation, compared to the evaluation of the actual trait, choosing a role-model based on previous choices made may be advantageous on its own. % TODO rephrase
This bias depends on the state of a role-model rather than on its trait, in contrast to frequency biases such as conformity,% TODO cite doi 10.1111/j.1420-9101.2011.02236.x
 which depend on the frequency of a trait in the population or in a sample of role-models. 
 We define a model of cultural transmission with prestige bias that combines both success and influence biases, provide analytic approximations for this model, and analyze its dynamics.

@@

After finding that the DM distribution is a good approximation of the (within-generation) role-model choice process, we turn our attention to the (between-generation) evolutionary dynamics.
We focus on the fixation probability and fixation time of an advantageous phenotype, using a diffusion-equation approximation approach, similar to analyses of population-genetic models \citep{kimura,kimura_average,otto_fixation}.
We are mainly interested in the effect of the bias weight, $\alpha$, which determines the relative effect of success and influence on prestige bias. % TODO ref eq 10?
For simplicity, we do not include role-model estimation error in this analysis. % TODO meaning epsilon=0?
As shown above, transmission in our model is approximately DM distributed (\cref{cor:dirichlet} and \cref{eq:copier_choose}).

@@

Similarly, we can use $1-\beta$ and $N_e$ in the population-genetics fixation time approximation given by \citet[eq.~8]{kimura_average}.\\

\begin{result}[Fixation time approximation]
\begin{equation}\label{eq:kimura_t}
T = \frac{1-\pi}{1-\beta}\int_0^x\frac{e^{2(1-\beta) \xi}-1}{\xi(1-\xi)}d\xi+ \frac{\pi}{1-\beta}\int_x^1\frac{1-e^{-2(1-\beta)(1-\xi)}}{\xi(1-\xi)}d\xi
\end{equation}
% TODO double check this, because eq 17 in Kimura 1969 looks different.
% in particular, I don't see Ne here...
\end{result}
Note that these integrals cannot be solved in closed form, so we can only estimate them numerically.

\paragraph{Numerical validation.}
To validate our analysis we performed multiple simulations comparing our binary model with the classic Wright-Fisher model, using different $\alpha$ and $\beta$ each time, and using the corresponding values of $s=1-\beta$ and $N_e=\alpha N + (1-\alpha)$ for the Wright-Fisher simulations.
%In \cref{fig:var_alpha} we changed $\alpha$ (and $N_e$ accordingly) and used a constant $\beta$.
%In \cref{fig:var_beta} we changed $\beta$ and used a constant $\alpha$.
We find see that the two models have similar dynamics, and both are well approximated the the approximations in \cref{eq:kimura_p,eq:kimura_t}.

\begin{figure}[h]
% TODO 
% capital letters in panel labels
% bigger font
    \includegraphics[width=\linewidth]{../figures/binary/kimura_var_alpha.pdf}
  \caption{\textbf{Fixation probability and time approximation for varying bias weight.}
  Fixation probability (\textbf{a}) and time (\textbf{b}) for varying bias weight on the bottom x-axis. The corresponds effective population size is on the top x-axis.
  The approximation (black) fits DM simulations results (green) and Wright-Fisher simulation results (orange) when using \cref{eq:kimura_p}.
  Fixation probability (a) is limited at approximately $2s$ (blue), as in the classic WF model.
  Markers are averages of $5,000$ simulations, error bars show 95\% confidence intervals.
   Here, Population size, $N=1,000$; phenotype values, $\hat{A}=1$, $A=0.7$; success coefficient, $1-\beta=s=0.044$.}
  \label{fig:var_alpha}
\end{figure}

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/binary/kimura_var_beta.pdf}
  \caption{\textbf{Fixation probability and time approximation for varying success coefficient.}
  % TODO
  % merge with previous figure; see comments there.
   The approximation (black) fits DM simulations results (green) and Wright Fisher simulation results (orange) when using \cref{eq:kimura_p}.
  Markers for average values across $5,000$ simulations, error bars for 75\% confidence intervals.
 Here, Population size, $N=1,000$; phenotype values, $\hat{A}=1$, $A=a\cdot \hat{A}, a \in [0.01,0.99]$; bias weight, $\alpha=0.01$.}
  \label{fig:var_beta}
\end{figure}

@@

\begin{result} [Drift and diffusion terms in a changing environment]\label{res:ch_expected}
Let $x$ be the initial frequency of the invading phenotype and $X_t$ is the number of individuals with the phenotype at time $t$.
Then,
\begin{equation}\label{eq:ch_expected_and_var}
E[X_t/N - x] \simeq x(1-x) S_t / N \;, 
\quad
V(X_t/N ) \simeq x(1-x) t/N_e \;, % TODO shouldn't it be V[X_t/N - x] ?? or something like that? with square brackets and with "- something" ?
\end{equation}
where $S_t=\sum\limits_{i=1}^{t} N (1-\beta_t)$.
\end{result}

@@

Using \cref{eq:ch_1} we see that
\begin{equation}\label{eq:ch_var1}
% TODO check this - looks weird that you have such a similar expression on the left side
\begin{split}
&E\left[\frac{X_{t+1}}{N} \bigg|X_t \right] - E\left[\frac{X_{t}}{N} \bigg|X_t \right] 
= \frac{1}{N}s_{t+1}\frac{X_t}{N}\left(1-\frac{X_t}{N} \right) \\
&E\left[\frac{X_{t+1}}{N} \bigg|X_t \right] 
= \frac{X_t}{N} + \frac{1}{N}s_{t+1}\frac{X_t}{N}\left(1-\frac{X_t}{N} \right) \;.
\end{split}
\end{equation}

@@

Using the drift and diffusion terms and following \citet{changeEnv}, we can approximate the fixation probability in a changing environment using
\begin{equation}\label{eq:ch_env}
\tilde\pi = \frac{1-e^{-2 \frac{S_n}{n} N_e x}}{1-e^{-2 \frac{S_n}{n} N_e}}
\end{equation}
where $\frac{S_n}{n} = \frac{k-l}{k+l}(1-\beta)$ and $n=k+l$.
That is, we use the average selection coefficient during a cycle of $k+l$ generations as the selection coefficient \cref{eq:kimura_p}.
% TODO what are you saying in the next paragraph?
%In our proof we showed that the expected change in frequency and variance is only manifested in the selection coefficient $S_n$, and that we can use those modified equation as a base for Kimura's equation.

\paragraph{Numerical validation.}
Comparing our approximation (\cref{eq:ch_env}) to numerical simulations, we find that the approximation fits simulations results well for variable bias weights, $\alpha$, which corresponds to the effective population size (\Cref{fig:ch_env_alpha_beta}a).

However, the approximation is more sensitive to the value of the success bias coefficient $\beta$ (\Cref{fig:ch_env_alpha_beta}b).
We suspect that when $\beta$ is too small, there will not be many cycles in the simulations, because either the population reaches a high frequency of the fitter phenotype after just a few cycles, or the fitter phenotype becomes extinct very quickly. 
For the very low $\beta$ values (0.65 and below),% TODO is that really "very low?"
the fixation probability exceeds even the constant environment approximation (which is the upper limit). However, this is to be expected, because the diffusion-equation approximations assume weak selection (i.e., low selection coefficient $s$).

% TODO what is k and l?
We found that for large $k$-to-$l$ ratio (while keeping constant total cycle length, $n=k+l=40$) the changing environment approximation (\cref{eq:ch_env}) converges to the constant environment approximation (\cref{eq:p_kimura}), see \Cref{fig:ch_env_k_l}. 
When using higher values of $n$, the agreement between the approximation and simulation results is weaker. % TODO reference to figure?
This is because our approximation is more precise when more cycles occur in the simulations. % TODO "in the simulations"-right?
When $n$ is high, the cycles are longer, the dynamics are therefore similar to a constant environment, and fixation or extinction is more likely to happen before the next cycle. Thus, there will be less cycles, which lowers our approximation precision.

\begin{figure}[h]
% TODO
% See comments on previous figures
% make x axes as before, with alpha on bottom and Ne on top
    \includegraphics[width=\linewidth]{../figures/changed_env/ch_env_var.pdf}
  \caption{\textbf{Sensitivity of fixation probability approximation in a changing environment to varying success coefficient and bias weight.}
\textbf{)a)} When the bias weight, or effective population size varies, the approximation (blue; \cref{eq:ch_env}) fits simulation results (orange). 
\textbf{(b)} When success bias is large ($\beta > 0.1$),  % TODO it was s>0.1, and O changed to beta>0.1 because that's what the x axes says, but I am not sure, as below you say the approximation works for weak selection, and s=1-beta, so large beta should give weak selection... please check
simulation results (orange) are underestimated by the approximation (blue; \cref{eq:ch_env}). With even larger success bias ($\beta > 0.35$), even the constant environment approximation (green; \cref{eq:kimura_p}) underestimate simulation results, likely because the diffusion equation approximation assumes weak selection.
Markers show average of $10,000$ simulations, error bars show are 75\% (a) and 95\% (b) confidence intervals.
  Here, population size, $N=1,000$; phenotype values, $\hat{A}=1$, $A=0.9$; \textbf{(a)} success bias/selection coefficient is: $1-\beta=s=0.005$, \textbf{(b)} success bias weight is: $\alpha=0.1$. 
  }
  \label{fig:ch_env_alpha_beta}
\end{figure}

\begin{figure}[h]
% TODO
% merge with previous figure and edit the caption accordingly.
    \includegraphics[width=\linewidth]{../figures/changed_env/ch_env_var_k_l.pdf}
  \caption{\textbf{Robustness of approximation in a changing environment to varying sizes of the cycles.} 
Markers for average values across $10,000$ simulations, error bars are 95\% confidence intervals.
The approximation (blue) fits the simulations (orange) well for different sizes of the changing environment cycle, where $k$ is the number of generations where the bias is advantageous, and $l$ are the generations it is disadvantageous.
   When $k<l$ the approximation is good. When $k>l$, the approximation and the simulations are both very close to the constant environment approximation (green). 
 Here, population size, $N=1,000$; phenotype values, $\hat{A}=1$, $A=0.8$; success coefficient, $1-\beta=s=0.02$; success bias weight, $\alpha=0.1$.}
  \label{fig:ch_env_k_l}
\end{figure}

@@

We ran simulations of the role-model choice process during a single generation in which every copier evaluates its own optimal $\alpha$ value. This is the $\alpha$ value that minimizes the mean error in estimation. % TODO estimation of what?
This is computed by minimizing the mean difference between the estimated value of every role-model and the ideal trait value $\hat{A}$. % TODO write the explicit equation

We see in \cref{fig:influence_advantage,fig:const_influence} a comparison between a model where every copier chooses its ideal success bias weight $\alpha_{ij}$, and when $\alpha$ is a constant homogeneous value.
We find that when copiers adapt their success-bias weight, their success-bias weight decreases with the number of copiers that have already chose a role-model~(\cref{fig:influence_advantage}a).
Moreover, their estimation error is much lower compared to a constant success-bias weight, which gives roughly the same high estimation error to all copiers (compare \cref{fig:influence_advantage}b and c): the estimation error converges to $0.046$ whereas a constant weight gives values $>0.74$ in this example.

\begin{figure}[b!]
    \includegraphics[width=\linewidth]{../figures/influence_advantage/choose_bias.pdf}
    % TODO
    % fonts too small
    % capital letters in labels
    % add "alpha" to ylabel in panel a
    % "mean estimation error" in y label of panel b
	% can you maybe add two more lines with different colors on the same plots for other parameters choices? say larger pop size or larger error?
  \caption{
  \textbf{Advantage of an adaptive success-bias weight.}
  Both success-bias weight $\alpha$ (\textbf{a}) and estimation error (\textbf{b}) 
   decrease during the role-model choosing process, demonstrating that influence is becomes more advantageous as more copiers have made their choice.
   Here, population size $N=200$, estimation error is normally distributed $e \sim N(0,\eta^2)$, standard deviation $\eta=0.0005$, plots are average of $300$ simulations.}	
  \label{fig:influence_advantage}
\end{figure}

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/influence_advantage/const_bias.pdf}
    % TODO 
    % merge with previous plot
  \caption{
  \textbf{Estimation error when using constant success bias weight.}
  The mean estimation error doesn't have any significant improvement relative to when every copier chooses a bias for himself.
  The minimum error is no less than $0.74$, while the mean error of the model where $\alpha$ is chosen is $0.06$, converging to $0.046$, which is a factor of $10$ less than the constant $\alpha$ model, showing the clear advantage of choosing $\alpha$.
  Here, population size $N=200$, estimation error is normally distributed $e \sim N(0,\eta^2)$, standard deviation $\eta=0.0005$, plots are average of $300$ simulations.}	
  \label{fig:const_influence}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
\paragraph{Summary.}
During cultural transmission, cultural traits such as attitudes, values, beliefs, and behavioral patterns are transmitted between individuals, for example via copying and social learning.
Some cultural traits or cultural role-models may be more likely to be copied due to transmission biases. 
A common one is success bias, in which copiers are more likely to copy a successful role-model. Many models assume that success can be accurately estimated.
However, it has been suggested \citep{complexityPaper} that because estimating success is hard, a more common bias is \textit{prestige bias}---a bias towards role-models perceived to be successful, portrayed by other traits or means.

@@

In addition, we see that the process of choosing itself is necessary, and when using a constant ratio ($\alpha$), the mean estimation error in the population is greater by a factor of $10$, even when using the $\alpha$ the population converged to when being able to choose it (\cref{fig:const_influence}).
After we showing that choosing one's success bias weight is beneficial to the copier, we continued to create our models- continuous and dichotomous, that incorporated both success and influence. However, running simulations for a nested iterative model is costly, and analyzing such model mathematically is very hard. We therefor found approximations for the role-model choice process: the generalized binomial distribution (\cref{res:GBD}) and the Dirichlet-Multinomial distribution (\cref{cor:dirichlet}).
To use approximate our model with the GBD distribution, we must assume homogeneous estimation error, $e$, and homogeneous success bias weight, $\alpha$. Both assumptions are necessary to this approximation, because it requires all copiers to be identical (order of copiers must not be a factor).
For the DM approximation we can relax the assumption for $\alpha$, and allow a varying $\alpha_j$ for each role-model $j$. This means that a copier is still not able to choose a success bias weight for itself, and instead it is a characteristic of the role-model only.
Using the DM approximation, because it is easier to analyze, we found equivalents for Kimura's equations, for our dichotomous model.
Following \cref{res:selection_coef,res:effective_population} we found an approximation for the fixation probability of the ideal trait, and an approximation for the time of fixation. These approximations hold only when the equivalent for the selection coefficient, in our model $1-\beta(A)$, is low (0.4 or lower). We also discovered that changing the success bias weight doesn't hurt the goodness of the approximations, and only affects the effective population size $N_e$.
Once we found the equations for a constant environment, we moved on to a cyclic changing environment.
Based on \cref{res:ch_expected,res:ch_var}, we found the approximations for the fixation probability and time to fixation.
We discovered that as in the constant environment, the effective population size doesn't affect the goodness of fit of the approximation. The success coefficient however, must be even lower than before for the approximation to fit. When the success coefficient is higher than 0.15, the simulation results were located above the changing environment approximation, and below the constant environment approximation. We believe the reason is the structure of the cycle.
Our proof and approximation in the changing environment are for a large amount of cycles, and when the success coefficient is too high, there might be very few cycles. Either the ideal trait is copied by enough copiers so that the influence is sufficient to negate the success bias when the cycle changes (and the beneficial trait becomes the maladaptive), or the opposite happens, and the ideal trait gets extinct before there were enough copiers that copied it.
We then tried to change the ratio between the number of cycles where $\hat{A}$ is beneficial and maladaptive. We showed that the approximation fits well regardless of the ratio, but when the ratio of adaptive generations to maladaptive ones is very high, it is very similar to a constant environment model.

>>>>

\paragraph{Influence bias.}
Today, social media provides an easy way to estimate the social and cultural influence individuals have over others, and therefore may have an effect on decision making. Online social networks such as \textit{Facebook} or \textit{Instagram} are known to affect the influence of individuals \citep{social_influence,influence_analysis,social_media}, and specific marketing practices were invented to capitalize on this effect \citep{facebook_marketing}.
Here, we model such influence as an indirect bias in cultural transmission, in which the choice of a role-model depends on the choices made by other individuals that have already chosen a role-model. 
In addition to using influence bias as an easy way to approximate role-models worth, it can also be a used to approximate it better than the copier's own estimation. This bias can save the copier costs: time, effort etc, and it can also improve its choice of role-model.

This bias depends on the state of a role-model rather than on its trait, in contrast to frequency biases such as conformity \citep{kin_selection}, which depend on the frequency of a trait in the population or in a sample of role-models. 
 We define a model of cultural transmission with prestige bias that combines both success and influence biases, provide analytic approximations for this model, and analyze its dynamics.

@@

After finding that the DM distribution is a good approximation of the (within-generation) role-model choice process, we turn our attention to the (between-generation) evolutionary dynamics.
We focus on the fixation probability and fixation time of an advantageous phenotype, using a diffusion-equation approximation approach, similar to analyses of population-genetic models \citep{kimura,kimura_average,otto_fixation}.
We are mainly interested in the effect of the bias weight, $\alpha$, which determines the relative effect of success and influence on prestige bias, given by \cref{eq:prestige}.
For simplicity, we do not include role-model estimation error in this analysis, i.e $e_i=0$ for every copier $i$.
As shown above, transmission in our model is approximately DM distributed (\cref{cor:dirichlet} and \cref{eq:copier_choose}).

@@

Similarly, we can use $1-\beta$ and $N_e$ in the population-genetics fixation time approximation given by \citet[eq.~17]{kimura_average}.\\

\begin{result}[Fixation time approximation]
\begin{equation}
T = J_1 + \frac{1-u(x)}{u(x)}\cdot J_2,
\end{equation}
where, given $S=Ne(1-\beta)$ and $x=\frac{1}{N_e}$,
\begin{equation}
u(x)=\frac{1-e^{-2Sx}}{1-e^{2S}}
\end{equation}
and,
\begin{equation}
J_1 = \frac{2}{(1-\beta)(1-e^{-2S})}\int_x^1 \frac{(e^{2S\xi}-1)(e^{-2S\xi}-e^{-2S})}{\xi(1-\xi)}d\xi,
\end{equation}
\begin{equation}
J_2= \frac{2}{(1-\beta)(1-e^{-2S})}\int_0^p \frac{(e^{2S\xi}-1)(1-e^{-2S\xi})}{\xi(1-\xi)}d\xi
\end{equation}

\begin{equation}\label{eq:kimura_t}
T = \frac{1-\pi}{1-\beta}\int_0^x\frac{e^{2(1-\beta) \xi}-1}{\xi(1-\xi)}d\xi+ \frac{\pi}{1-\beta}\int_x^1\frac{1-e^{-2(1-\beta)(1-\xi)}}{\xi(1-\xi)}d\xi
\end{equation}
\end{result}
Note that these integrals cannot be solved in closed form, so we can only estimate them numerically.

\paragraph{Numerical validation.}
To validate our analysis we performed multiple simulations comparing our binary model with the classic Wright-Fisher model, using different $\alpha$ and $\beta$ each time, and using the corresponding values of $s=1-\beta$ and $N_e=\alpha N + (1-\alpha)$ for the Wright-Fisher simulations.

We find see that the two models have similar dynamics, and both are well approximated the the approximations in \cref{eq:kimura_p,eq:kimura_t}.

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/binary/kimura_var.pdf}
  \caption{\textbf{Fixation probability and time approximation for varying bias weight and success coefficient.}
  Fixation probability (\textbf{A}) and time (\textbf{B}) for varying bias weight on the bottom x-axis. The corresponds effective population size is on the top x-axis.
  The approximation (black) fits DM simulations results (green) and Wright-Fisher simulation results (orange) when using \cref{eq:kimura_p}.
  Fixation probability (\textbf{A}) is limited at approximately $2s$ (blue), as in the classic WF model.
  Markers are averages of $10,000$ simulations, error bars show 95\% confidence intervals for (\textbf{A,B}) and 75\% for (\textbf{C,D}).
   Here, Population size, $N=1,000$; phenotype values, $\hat{A}=1$, $A=0.7$ (\textbf{A,B}), $A=a\cdot \hat{A}, a \in [0.01,0.99]$ (\textbf{C,D}); success coefficient, $1-\beta=s=0.044$ (\textbf{A,B}); bias weight, $\alpha=0.01$ (\textbf{C,D}).}
  \label{fig:var_alpha}
\end{figure}

@@

\begin{result} [Drift and diffusion terms in a changing environment]\label{res:ch_expected}
Let $x$ be the initial frequency of the invading phenotype and $X_t$ is the number of individuals with the phenotype at time $t$.
Then,
\begin{equation}\label{eq:ch_expected_and_var}
E[X_t/N - x] \simeq x(1-x) S_t / N \;, 
\quad
V(X_t/N-x) = V(X_t/N) \simeq x(1-x) t/N_e \;,
\end{equation}
where $S_t=\sum\limits_{i=1}^{t} N (1-\beta_t)$.
\end{result}

@@

Using \cref{eq:ch_1} we see that
\begin{equation}\label{eq:ch_var1}
\begin{split}
&E\left[\frac{X_{t+1}}{N} \bigg|X_t \right] - E\left[\frac{X_{t}}{N} \bigg|X_t \right] 
= \frac{1}{N}s_{t+1}\frac{X_t}{N}\left(1-\frac{X_t}{N} \right) \\
&E\left[\frac{X_{t+1}}{N} \bigg|X_t \right] 
= \frac{X_t}{N} + \frac{1}{N}s_{t+1}\frac{X_t}{N}\left(1-\frac{X_t}{N} \right) \;.
\end{split}
\end{equation}

@@

Using the drift and diffusion terms and following \citet{changeEnv}, we can approximate the fixation probability in a changing environment using
\begin{equation}\label{eq:ch_env}
\tilde\pi = \frac{1-e^{-2 \frac{S_n}{n} N_e x}}{1-e^{-2 \frac{S_n}{n} N_e}}
\end{equation}
where $\frac{S_n}{n} = \frac{k-l}{k+l}(1-\beta)$ and $n=k+l$.
That is, we use the average selection coefficient during a cycle of $k+l$ generations as the selection coefficient \cref{eq:kimura_p}.

\paragraph{Numerical validation.}
Comparing our approximation (\cref{eq:ch_env}) to numerical simulations, we find that the approximation fits simulations results well for variable bias weights, $\alpha$, which corresponds to the effective population size (\Cref{fig:ch_env_alpha_beta}A).

However, the approximation is more sensitive to the value of the success bias coefficient $\beta$ (\Cref{fig:ch_env_alpha_beta}B).
We suspect that when $\beta$ is too small, there will not be many cycles in the simulations, because either the population reaches a high frequency of the fitter phenotype after just a few cycles, or the fitter phenotype becomes extinct very quickly. 
For the $\beta$ values (0.65 and below), the fixation probability exceeds even the constant environment approximation (which is the upper limit). However, this is to be expected, because the diffusion-equation approximations assume weak selection (i.e., low selection coefficient $s$).

When $k$ is the number of generations where the invading phenotype is advantageous, and $l$ when the original trait is, we found that for large $k$-to-$l$ ratio (while keeping constant total cycle length, $n=k+l=100$), the changing environment approximation (\cref{eq:ch_env}) converges to the constant environment approximation (\cref{eq:kimura_p}), see \Cref{fig:ch_env_alpha_beta} (C,D).

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/changed_env/ch_env.pdf}
  \caption{\textbf{Sensitivity of fixation probability approximation in a changing environment to varying success coefficient and bias weight and robustness to varying ratios of cycles.}
\textbf{(A)} When the bias weight, or effective population size varies, the approximation (blue; \cref{eq:ch_env}) fits simulation results (orange). 
\textbf{(B)} When success bias is large ($1-\beta > 0.1$),  
simulation results (orange) are underestimated by the approximation (blue; \cref{eq:ch_env}). With even larger success bias ($1-\beta > 0.35$), even the constant environment approximation (green; \cref{eq:kimura_p}) underestimate simulation results, likely because the diffusion equation approximation assumes weak selection.
We can also see the robustness of the approximation to changes in cycle ratios (\textbf{C,D}): The approximation (blue) fits the simulations (orange) well for different sizes of the changing environment cycle, where $k$ is the number of generations where the bias is advantageous, and $l$ are the generations it is disadvantageous.
When $k<l$ the approximation is good. When $k>l$, the approximation and the simulations are both very close to the constant environment approximation (green). 
Markers show average of $10,000$ simulations, error bars show are 75\% \textbf{(A,C,D)} and 95\% \textbf{(B)} confidence intervals.
  Here, population size, $N=1,000$; phenotype values, $\hat{A}=1$, $A=0.9$ (\textbf{A,B}), $A=0.8$ \textbf{(C,D)}; In \textbf{(A)} success bias/selection coefficient is: $1-\beta=s=0.005$; In \textbf{(B,C,D)} success bias weight is: $\alpha=0.1$.}
  \label{fig:ch_env_alpha_beta}
\end{figure}

@@

We ran simulations of the role-model choice process during a single generation in which every copier evaluates its own optimal $\alpha$ value. This is the $\alpha$ value that minimizes the mean square error of the difference between the estimated and the ideal trait values.
This $\alpha$ is found by minimizing the following:
\begin{equation}
SMSE = \sum_{j=1}^N\frac{\alpha A_j + (1-\alpha) K_j}{\sum_{l=1}^N\alpha A_l + (1-\alpha) K_l}\cdot (\hat{A}-A_j)^2
\end{equation}
where $A_j$ is the trait of role-model $j$ and $K_j$ the number of copiers that already chose him.

We see in \cref{fig:influence_advantage} a comparison between a model where every copier chooses its ideal success bias weight $\alpha_{ij}$, and when $\alpha$ is a constant homogeneous value.
We find that when copiers adapt their success-bias weight, their success-bias weight decreases with the number of copiers that have already chose a role-model~(\cref{fig:influence_advantage}).
Moreover, their estimation error is much lower compared to a constant success-bias weight, which gives roughly the same high estimation error to all copiers (compare \cref{fig:influence_advantage}b and c): the estimation error converges to $0.046$ whereas a constant weight gives values $>0.74$ in this example.

\begin{figure}[b!]
    \includegraphics[width=\linewidth]{../figures/influence_advantage/choose_bias.pdf}
  \caption{
  \textbf{Advantage of an adaptive success-bias weight.}
  Both success-bias weight $\alpha$ (\textbf{a}) and estimation error (\textbf{b}) decrease during the role-model choosing process, demonstrating that influence is becomes more advantageous as more copiers have made their choice.
However, when $\alpha$ is homogeneous, the mean estimation error doesn't decrease, regardless of $\alpha$ or $\eta$.
The mean estimation error in the homogeneous $\alpha$ model is larger by a factor of $10$ than the adaptive $\alpha$ model.
     Here, population size $N=200$; estimation error is normally distributed $e \sim N(0,\eta^2)$; standard deviation $\eta=0.0001 (blue), 0.001 (orange), 0.01 (green)$, plots are average of $300$ simulations.}	
  \label{fig:influence_advantage}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\section*{Discussion}
\paragraph{Summary.}
During cultural transmission, cultural traits such as attitudes, values, beliefs, and behavioral patterns are transmitted between individuals, for example via copying and social learning.
Some cultural traits or cultural role-models may be more likely to be copied due to transmission biases. 
A common one is success bias, in which copiers are more likely to copy a successful role-model. Many models assume that success can be accurately estimated.
However, it has been suggested \citep{complexityPaper} that because estimating success is hard, a more common bias is \textit{prestige bias}---a bias towards role-models perceived to be successful, portrayed by other traits or means.

@@

In addition, we see that the process of choosing itself is necessary, and when using a constant ratio ($\alpha$), the mean estimation error in the population is greater by a factor of $10$, even when using the $\alpha$ the population converged to when being able to choose it (\cref{fig:influence_advantage}).
After we showing that choosing one's success bias weight is beneficial to the copier, we continued to create our models- continuous and dichotomous, that incorporated both success and influence. However, running simulations for a nested iterative model is costly, and analyzing such model mathematically is very hard. We therefor found approximations for the role-model choice process: the generalized binomial distribution (\cref{res:GBD}) and the Dirichlet-Multinomial distribution (\cref{cor:dirichlet}).
To use approximate our model with the GBD distribution, we must assume homogeneous estimation error, $e$, and homogeneous success bias weight, $\alpha$. Both assumptions are necessary to this approximation, because it requires all copiers to be identical (order of copiers must not be a factor).
For the DM approximation we can relax the assumption for $\alpha$, and allow a varying $\alpha_j$ for each role-model $j$. This means that a copier is still not able to choose a success bias weight for itself, and instead it is a characteristic of the role-model only.
Using the DM approximation, because it is easier to analyze, we found equivalents for Kimura's equations, for our dichotomous model.
Following \cref{res:selection_coef} we found an approximation for the fixation probability of the ideal trait, and an approximation for the time of fixation. These approximations hold only when the equivalent for the selection coefficient, in our model $1-\beta(A)$, is low (0.4 or lower). We also discovered that changing the success bias weight doesn't hurt the goodness of the approximations, and only affects the effective population size $N_e$.
Once we found the equations for a constant environment, we moved on to a cyclic changing environment.
Based on \cref{res:ch_expected}, we found the approximations for the fixation probability and time to fixation.
We discovered that as in the constant environment, the effective population size doesn't affect the goodness of fit of the approximation. The success coefficient however, must be even lower than before for the approximation to fit. When the success coefficient is higher than 0.15, the simulation results were located above the changing environment approximation, and below the constant environment approximation. We believe the reason is the structure of the cycle.
Our proof and approximation in the changing environment are for a large amount of cycles, and when the success coefficient is too high, there might be very few cycles. Either the ideal trait is copied by enough copiers so that the influence is sufficient to negate the success bias when the cycle changes (and the beneficial trait becomes the maladaptive), or the opposite happens, and the ideal trait gets extinct before there were enough copiers that copied it.
We then tried to change the ratio between the number of cycles where $\hat{A}$ is beneficial and maladaptive. We showed that the approximation fits well regardless of the ratio, but when the ratio of adaptive generations to maladaptive ones is very high, it is very similar to a constant environment model.
